### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/manual_axisem1.2.tex

%USER MANUAL%
\color{dark_grey}
\hfill{\Huge \fontfamily{\sfdefault}\selectfont User Manual \\
% FILL: manual version
% e.g. 1.0
\raggedleft \huge \fontfamily{\sfdefault}\selectfont Version {1.1}\\}

@@

\begin{center}
    \begin{tabularx}{\textwidth}{lX}
        \begin{minipage}{0.3\textwidth}
            \begin{center}
                \includegraphics[width=\textwidth]{Axisem2png.png}
            \end{center}
        \end{minipage}
        &
        \begin{minipage}{0.6\textwidth}
            \begin{center}
                \title{}
                \LARGE{ \textbf{\sc AxiSEMv1.1 Manual}}
                \vspace*{0.6cm}\\
                {\large 
                Tarje Nissen-Meyer\textsuperscript{1},
                Martin van Driel\textsuperscript{2},
                Simon St\"{a}hler\textsuperscript{3},
                Kasra Hosseini\textsuperscript{3},
                Stefanie Hempel\textsuperscript{4}, 
                Alexandre Fournier\textsuperscript{5}}
                \vspace*{0.3cm}\\
                {\small  \textsuperscript{1} Oxford University (UK),
                  \textsuperscript{2} ETH Zurich (Switzerland),  \textsuperscript{3}
                  LMU M\"{u}nchen (Germany), \textsuperscript{4} Universit\"{a}t
                  M\"{u}nster (Germany), \textsuperscript{5} IPG Paris
                  (France)}\\
                {\large Oxford, UK -- July 7, 2014}
            \end{center}
        \end{minipage}
    \end{tabularx}
\end{center}

@@

\subsubsection{NetCDF}

AxiSEM allows to output larger datasets, especially wavefields in the NetCDF format. The
current version also has full support for binary dumps, but the development will move
towards NetCDF containers.\\
Unfortunately, the installation of the NetCDF libraries may still be cumbersome; below we list some limited support for installing NetCDF..

@@

\begin{center}
    %\includegraphics[height=50mm]{mesher_comp_cost.pdf}
    \includegraphics[height=50mm]{memory_usage_new.pdf}
    \hspace{5mm}
    \includegraphics[height=50mm]{solver_totalcpu.pdf}
\end{center}

On the left, you may deduct the mesher's RAM occupation as a function of frequency. Going
towards very high resolution (around and above 1Hz), you will need a rather fat node (>
16GB RAM) for the (shared memory) meshing.
On the right, we depict the computational cost
associated with the solver to compute seismograms of 30 min length. The relation between
seismic period and CPU-hours is an approximate proxy to estimate how many cores and
wall-clock time is optimal for your infrastructure. Scaling of CPU hours is approximately 
to the third power of the maximum frequency.

@@

\verb|SOLVER/inparam_source|, \verb|SOLVER/CMTSOLUTION| and \verb|SOLVER/inparam_basic|.
AxiSEM has two principal modes, which are selected by the value \verb|SIMULATION_TYPE| in
\verb|SOLVER/inparam_basic|.

@@

\item \verb|SIMULATION_TYPE moment|:\\
      The submit.csh script starts four separate simulations for the basic types $M_{rr}$,
      $M_{\theta\theta}+M_{\phi\phi}$, $M_{\theta r}$, $M_{\theta \phi}$. You have to run
      the postprocessing script after the simulation to sum them up correctly. \\ Before
      the simulation, set the source depth and the moment tensor in
      \verb|SOLVER/CMTSOLUTION|. \\Run \verb|postprocessing.csh| in the simulation
      directory afterwards. You can run postprocessing for different moment tensors on the
      same simulation, but not for different depths (since the forward simulation depends
      on the depth).  The \verb|CMTSOLUTION| file is a standard format and can be
      downloaded from many sites in the web, including
      \url{http://www.globalcmt.org/CMTsearch.html}
\end{enumerate}

@@

@@

(5) Tarje Nissen-Meyer (2007),
\textit{Full-wave seismic sensitivity in a spherical Earth},
Ph.D. thesis, Princeton University
(This includes refs (2)-(4) and more details.)\\
% 
% (5) Jean-Paul Ampuero, Tarje Nissen-Meyer (2011),
% \textit{High-order conservative time schemes in spectral-element methods 
% for seismic wave propagation.}, To be submitted to Geophys. J. Int.\\

@@

(6) Deville, M. O., Fischer, P. F., Mund, E. H. (2002), 
\textit{High-Order Methods for Incompressible Fluid Flow}, 
Vol. 2, Cambridge monographs on Sppl. \& Comp. Math., Cambridge University Press.\\

(7) Tufo, H. M., Fischer, P. F. (2001), \textit{Fast Parallel Direct Solvers For Coarse
Grid Problems}, 61, 151-177, J. Par. and Dist. Comput.\\

(8) Bernardi, C., Dauge, M., Maday, Y. (1999), \textit{Spectral Methods for Axisymmetric
Domains}, Vol. 3, Series in Appl. Math., Gauthier-Villars, Paris.\\

(9) Chaljub, E. (2000), \textit{Mod{\'{e}}lisation num{\'{e}}rique de la 
propagation d'ondes sismiques en g{\'{e}}om{\'{e}}trie sph{\'{e}}rique:
Application {\`{a}} la sismologie globale}, 
Ph.D. thesis, Universit{\'{e}} de Paris 7.\\

(10) Komatitsch D., Tromp, J. (2002), \textit{Spectral-element simulations of
global seismic wave propagation---I. Validation},
149, 390-412, Geophys. J. Int.

>>>>

%USER MANUAL%
\color{dark_grey}
\hfill{\Huge \fontfamily{\sfdefault}\selectfont User Manual \\
% FILL: manual version
% e.g. 1.0
\raggedleft \huge \fontfamily{\sfdefault}\selectfont Version {1.2}\\}

@@

\begin{center}
    \begin{tabularx}{\textwidth}{lX}
        \begin{minipage}{0.3\textwidth}
            \begin{center}
                \includegraphics[width=\textwidth]{Axisem2png.png}
            \end{center}
        \end{minipage}
        &
        \begin{minipage}{0.6\textwidth}
            \begin{center}
                \title{}
                \LARGE{ \textbf{\sc AxiSEMv1.2 Manual}}
                \vspace*{0.6cm}\\
                {\large 
                Tarje Nissen-Meyer\textsuperscript{1},
                Martin van Driel\textsuperscript{2},
                Simon St\"{a}hler\textsuperscript{3},
                Kasra Hosseini\textsuperscript{3},
                Stefanie Hempel\textsuperscript{4}, 
                Alexandre Fournier\textsuperscript{5}}
                \vspace*{0.3cm}\\
                {\small  \textsuperscript{1} Oxford University (UK),
                  \textsuperscript{2} ETH Zurich (Switzerland),  \textsuperscript{3}
                  LMU M\"{u}nchen (Germany), \textsuperscript{4} Universit\"{a}t
                  M\"{u}nster (Germany), \textsuperscript{5} IPG Paris
                  (France)}\\
                {\large Oxford, UK -- July 7, 2014}
            \end{center}
        \end{minipage}
    \end{tabularx}
\end{center}

@@

\subsubsection{NetCDF}
\label{sssec:netcdf}

AxiSEM allows to output larger datasets, especially wavefields in the NetCDF format. \\
Unfortunately, the installation of the NetCDF libraries may still be cumbersome; below we
list some limited support for installing NetCDF..

@@

\begin{center}
    %\includegraphics[height=50mm]{mesher_comp_cost.pdf}
    \includegraphics[height=50mm]{memory_usage_new.pdf}
    \hspace{5mm}
    %\includegraphics[height=50mm]{solver_totalcpu.pdf}
    \includegraphics[height=50mm]{db_generation_cost.pdf}
\end{center}

On the left, you may deduct the mesher's RAM occupation as a function of frequency. Going
towards very high resolution (around and above 1Hz), you will need a rather fat node (>
16GB RAM) for the (shared memory) meshing.
On the right, we depict the computational cost
associated with the solver to compute seismograms of one hour length for Earth and Mars
and for the second order Newmark as well as the fourth order symplectic time scheme. The
relation between seismic period and CPU-hours is an approximate proxy to estimate how many
cores and wall-clock time is optimal for your infrastructure. Scaling of CPU hours is
approximately to the third power of the maximum frequency. As a rule of thumb, using the
symplectic scheme is advisable when propagating waves for more than 100 wavelengths.

@@

\verb|SOLVER/inparam_source|, \verb|SOLVER/CMTSOLUTION| and \verb|SOLVER/inparam_basic|.
AxiSEM has three principal modes, which are selected by the value \verb|SIMULATION_TYPE| in
\verb|SOLVER/inparam_basic|.

@@

\item \verb|SIMULATION_TYPE moment|:\\
      The submit.csh script starts four separate simulations for the basic types $M_{rr}$,
      $M_{\theta\theta}+M_{\phi\phi}$, $M_{\theta r}$, $M_{\theta \phi}$. You have to run
      the postprocessing script after the simulation to sum them up correctly. \\ Before
      the simulation, set the source depth and the moment tensor in
      \verb|SOLVER/CMTSOLUTION|. \\Run \verb|postprocessing.csh| in the simulation
      directory afterwards. You can run postprocessing for different moment tensors on the
      same simulation, but not for different depths (since the forward simulation depends
      on the depth).  The \verb|CMTSOLUTION| file is a standard format and can be
      downloaded from many sites in the web, including
      \url{http://www.globalcmt.org/CMTsearch.html}

\item \verb|SIMULATION_TYPE force|:\\
      The submit.csh script starts two separate simulations for the basic types $P_r$ and
      $P_\theta$. This mode is used to compute the databases for instaseis. Note, that the
      postprocessing does not handle these runs yet appropriately.
\end{enumerate}

@@

\subsection{Output wavefields in netcdf format needed for \textit{Instaseis}}

\begin{itemize}
    \item Compile with netcdf (see section~\ref{sssec:netcdf} for hints on the
    installation): in \verb|make_axisem.macros| set \verb|USE_NETCDF| to \verb|true|
    and if netcdf is not in your \verb|PATH|, also set \verb|NETCDF_PATH| to the location
    of your netcdf installation. If you have netcdf compiled with MPI and run on a
    supercomputer, using \verb|USE_PARALLEL_NETCDF true| might be beneficial for short
    period runs.
    \item In \verb|SOLVER/inparam_advanced| set \verb|USE_NETCDF| to \verb|true| to
    use netcdf output in the solver
    \item In \verb|SOLVER/inparam_advanced| set \verb|DATA_DIR| to some location where you
    have enough free space and in case you use parallel IO this should be on a parallel
    file system
    \item further settings in \verb|SOLVER/inparam_advanced| to generate a full database,
    that is all epicentral distances and source down to 700km and using the fourth order
    spatial scheme:

\begin{verbatim}
SOURCE_FUNCTION     gauss_0

KERNEL_WAVEFIELDS   true

KERNEL_DUMPTYPE     displ_only
KERNEL_SPP          4
KERNEL_SOURCE       igno

KERNEL_COLAT_MIN    0.
KERNEL_COLAT_MAX    180.

KERNEL_RMIN         5671.
KERNEL_RMAX         7000.
\end{verbatim}

    \item in \verb|SOLVER/inparam_basic| set \verb|SIMULATION_TYPE| to \verb|force| to
    simulate both the vertical and the horizontal database. To generate just one of the
    two, set it to \verb|single| and select either \verb|thetaforce| or \verb|vertforce|
    in \\\verb|SOLVER/inparam_source|

    \item once the AxiSEM runs are finished, go to the rundirectory and run
    \verb|./fieltransform.sh| (in case of single runs just run \verb|./fieltransform|
    directly). This rechunks the wavefields in the netcdf file to optimize them for
    reading as time series instead of snapshots.

\end{itemize}

@@

(5) Tarje Nissen-Meyer (2007),
\textit{Full-wave seismic sensitivity in a spherical Earth},
Ph.D. thesis, Princeton University
(This includes refs (2)-(4) and more details.)\\

(6) Martin van Driel and Tarje Nissen-Meyer (2014),
\textit{Seismic Wave Propagation in Fully Anisotropic Axisymmetric Media},
Geophysical Journal International 199 (2), 880–893.\\
doi:10.1093/gji/ggu269.\\

(7) Martin van Driel and Tarje Nissen-Meyer (2014),
\textit{Optimized Viscoelastic Wave Propagation for Weakly Dissipative Media},
Geophysical Journal International 199 (2) 1078–1093.\\
doi:10.1093/gji/ggu314.\\

%
% (5) Jean-Paul Ampuero, Tarje Nissen-Meyer (2011),
% \textit{High-order conservative time schemes in spectral-element methods 
% for seismic wave propagation.}, To be submitted to Geophys. J. Int.\\

@@

(8) Martin van Driel, Lion Krischer, Simon Stähler, Kasra Hosseini, and Tarje Nissen-Meyer (2015),
\textit{Instaseis: Instant Global Broadband Seismograms Based on a Waveform Database},
To be submitted to Solid Earth\\

(9) Deville, M. O., Fischer, P. F., Mund, E. H. (2002), 
\textit{High-Order Methods for Incompressible Fluid Flow}, 
Vol. 2, Cambridge monographs on Sppl. \& Comp. Math., Cambridge University Press.\\

(10) Tufo, H. M., Fischer, P. F. (2001), \textit{Fast Parallel Direct Solvers For Coarse
Grid Problems}, 61, 151-177, J. Par. and Dist. Comput.\\

(11) Bernardi, C., Dauge, M., Maday, Y. (1999), \textit{Spectral Methods for Axisymmetric
Domains}, Vol. 3, Series in Appl. Math., Gauthier-Villars, Paris.\\

(12) Chaljub, E. (2000), \textit{Mod{\'{e}}lisation num{\'{e}}rique de la 
propagation d'ondes sismiques en g{\'{e}}om{\'{e}}trie sph{\'{e}}rique:
Application {\`{a}} la sismologie globale}, 
Ph.D. thesis, Universit{\'{e}} de Paris 7.\\

(13) Komatitsch D., Tromp, J. (2002), \textit{Spectral-element simulations of
global seismic wave propagation---I. Validation},
149, 390-412, Geophys. J. Int.
